Unsupervised Detection and Tracking of Arbitrary Objects with Dependent Dirichlet Process Mixtures

### 2012-03-27 ltx/paper_short_2.tex
submitted to IJCV. files in submitted_ijcv_20120327 directory

\begin{abstract}
This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos. It is intended to reduce the need for detection and localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and image qualities. The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be easily and efficiently extracted from the regions in a video that represent objects. This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference in the model to accomplish detection and tracking. This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
\end{abstract}

@@

We begin by describing characteristics of the extracted data (Section~\ref{sec:dataextraction}), and giving the generalized form of the model (Section~\ref{sec:modeldefinition}). To implement this model, one must specify a data extraction procedure and distributions for representing objects, which may be chosen to allow for arbitrary object tracking or tailored to a specific object type for a given application. In our implementation, we extract data via a basic frame differencing procedure and specify distributions useful for representing arbitrary objects (Section~\ref{sec:modelspecification}). We describe inference algorithms for our model and show how the output of these algorithms can be interpreted as object localization and tracking results (Section~\ref{sec:inference}). Our implementation is demonstrated on multiple synthetic and benchmark datasets. Standard performance metric values are computed to quantify results on the benchmark datasets (Section~\ref{sec:experiments}). We compare our performance metrics with those yielded by specialized detection and tracking algorithms tailored to specific objects. Our results support our hypothesis that approaches combining simple data extraction and a powerful model can perform detection and tracking of arbitrary objects at a level comparable to state-of-the-art, object specific algorithms.

@@

Methods that discern between the foreground and background regions of a video allow for data to be extracted from the areas in each frame where objects exist. Frame differencing and background subtraction are two such methods. Both record locations that exhibit motion relative to the background. Often, background subtraction refers to methods that compare an image containing targets with an image of the background only or with some model of the background that is learned as the video progresses \cite{piccardi2004background}, while frame differencing refers to methods that compare pairs of consecutive images in a video \cite{zhang2001segmentation}. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking \citep{perez_2002}.

@@

Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos. Blob tracking, a basic method for carrying out this goal, has found success in videos where objects are easily isolated from the background and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}. Blob tracking methods, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}. To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}. Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects in videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}. Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions extracted by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and graph based methods that carry out a similar task using spectral clustering \cite{fragkiadaki2011detection}. Another approach uses a Gaussian mixture model to cluster data extracted from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new object tracks.

@@

We desire a data extraction procedure that yields observations of the form
\begin{equation}
\centering
\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
\end{equation}
where each $\bold{x}$ corresponds to a point within an image region where an object (or foreground element)  is believed to exist, $\bold{x}^{s} \in \mathbb{R}^{2}$ denotes the spatial location of this point, $\bold{x}^{c} \in S_{1} \times \ldots \times S_{V}$ denotes some collection of local image features in the vicinity of this point, and $t \in \{1, \ldots, T \}$ denotes the time index.

@@

We also extract features $\bold{x}^{c}$ that capture image information in the vicinity of each extracted pixel. Examples of possible features include color distributions, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations. In principle, we can extract any image features that may be used to characterize the appearance of objects. In our implementation, we choose to extract only color information in the vicinity of each pixel. Incorporating color features allows our method to infer a distribution over color for each detected object; this improves its ability to distingush between adjacent objects and track objects through occlusion. To add this information, we let $\bold{x}^{c}$ represent a $V$ dimensional discrete distribution over some aspect of color (such as the hue) in the pixel's immediate vicinity. Details on this vector and how it is computed are given in Section~\ref{sec:dataextractioninexperiments}. We refer to the components of this vector as the ``color counts'' of the pixel.

@@

Instead of incorporating the cluster size variable $m_{k,t}$ directly in the GPUDDPM model (as it was in the definition given by \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of new set of variables called deletion variables. We introduce a deletion variable $d_{i,t}$ for each observation $\bold{x}_{i,t}$, which denotes the time at which the observation is removed from its assigned cluster. At each time, cluster sizes $m_{k,t}$ can be reconstructed from all previous assignments and deletion variables by
\begin{equation}
\label{compute_clust_size}
m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
\end{equation}
where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise. Additionally, we can define the deletion variable $d_{i,t}$ to be $d_{i,t} = t + l_{i,t}$, where $l_{i,t}$ can be thought of as the lifespan of an assignment. From the definition of distribution D (given by \eqref{D_distro}), the lifespan can be shown to be distributed geometrically, and can be written as
\begin{equation}
\label{del_rho_form}
l_{i,t} | \rho  \sim  \rho(1 - \rho)^{l_{i,t}}
\end{equation}
where the parameter $\rho$ is the same as that in \eqref{D_distro}.

@@

\subsubsection{Sampling Assignment Variables $c_{i,t}$}
\label{sec:sample_assignments}

@@

\subsubsection{Sampling Cluster Parameters $\theta_{k, t}$}

The conjugacy of appearance model and transition kernel distributions allow us to easily sample from the posterior distribution over the cluster parameters, which we can write
\begin{align}
\begin{split}
P(\theta_{k,t} | \ldots ) = & \hspace{1mm} P(\bold{x}_{i,t} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})
\\ & \times P(\theta_{k,t} | z_{k,t,1:M}) \\
 = & \hspace{1mm} \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{k,t}, \Sigma_{k,t} | \boldsymbol{\mu}_{N}, k_{N}, v_{N}, \Lambda_{N}) \\
  & \times \mathcal{D}ir( \bold{p}_{k,t} | \bold{q}_{N})
\end{split}
\end{align}
Where the parameters in the above distribution are given when the observations $\bold{x}_{1:N_{t},t}$, auxiliary variables $z_{k,t-1:t,1:M}$ for cluster k at time $t-1$ and $t$, are taken to be the ``observations'' for the following Bayesian updates
\begin{eqnarray}
\label{bayesian_update_1}
\kappa_{N} &=& \kappa_{0} + N \\
\label{bayesian_update_2}
\nu_{N} &=& \nu_{0} + N \\
\label{bayesian_update_3}
\boldsymbol{\mu}_{N} &=& \frac{\kappa_{0}}{\kappa_{0}+N} \boldsymbol{\mu}_{0}  +  \frac{N}{\kappa_{0}+N} \overline{\bold{x}}^{s}\\
\Lambda_{N} &=& \Lambda_{0} + S_{\bold{x}^{s}}\\
\label{bayesian_update_4}
\bold{q}_{N} &=& \bold{q}_{0} + \sum_{i=1}^{N} \bold{x}_{i}^{c}
\end{eqnarray}
where $N$ is the number of observations, $\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{x}^{s}$ and $\bold{x}^{c}$ respectively denote the spatial and color features of the observations, and $\overline{\bold{x}}$ and $S_{\bold{x}}$ respectively denote the sample mean and sample covariance for the set of observations $\bold{x}$, defined in \eqref{samplemean} and \eqref{samplecov}.

@@

\subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
\label{sec:sample_aux_vars}

@@

\subsubsection{Sampling Deletion Variables $d_{i,t}$}
Sampling deletion variables could be performed in a manner similar to how we sample the assignment variables in Section~\ref{sec:sample_assignments}, but this may be computationally expensive due to the large number of possible deletion times. To remedy this, the MH algorithm may again be used to generate samples $d_{i,t}^{*}$ from the posterior distribution over possible deletion times, where we use the proposal distribution
\begin{align}
\begin{split}
l_{i,t}  &\sim  \text{Geo}(\rho)  \\
d_{i,t}^{*}  &= l_{i,t} + t + 1
\end{split}
\end{align}
where $l_{i,t}$ denotes the geometrically distributed ``lifetime'', and we accept or reject this sample using the process described in Section~\ref{sec:sample_aux_vars}.

@@

This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference. SMC inference operates in the orginal GPUDDPM formulation (Section~\ref{sec:gpuddpm}). The algorithm is shown in Algorithm~\ref{alg:SMC}. At each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots,$ $\theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t},$ $\ldots,$ $m_{K_{t},t}$. A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a new set of particles are sampled from the set of weighted particles via a resampling process. Within each time step, Gibbs sampling is used to generate the samples associated with each particle.

@@

\subsubsection{Proposal Distribution for Assignments $c_{i,t}$}
\label{sec:smc_proposal_1}

@@

Performance evaluation metrics, which provide a standardized way of quantifying the success of a detection and tracking procedure on a given video, have started to become consistently used in the past four years. The metrics presented in \cite{kasturi_2008} and used in \cite{ellis_2010, taj_2007, lee_2009} have become well established metrics for evaluating the performance of object detection and tracking in videos and have been adopted by the Video Analysis and Content Extraction (VACE) program and the Classification of Events, Activities, and Relationships (CLEAR) consortium, two large-scale efforts concerned with video tracking and interaction analysis. The two metrics used to quantify the experimental results in this study are known as the Sequence Frame Detection Accuracy (SFDA) and Average Tracking Accuracy (ATA). Details on how these metrics are defined and computed are given in Appendix~\ref{sec:performancemetricdetails}.

@@

The ground-truth authored by the ViPER tool took the form of bounding boxes denoting the spatial postion of each object at each time step. Consequentially, to find the spatial overlap between results and ground-truth, which is intrinsic to both metrics, a rectangular bounding box was needed per object per time step from the results of the algorithm. We took the maximal and minimal axially aligned values of the oval inferred by our algorithm (as described in Section~\ref{sec:inferencetotrackingresults}) to be the sides of a representative bounding box for a given object at a given frame.

@@

Frame differencing was used in all experiments to identify pixels exhibiting motion. 
For each pixel $\bold{x} = ( x_{1},$ $x_{2},$ $t )$ recorded during frame differencing, we also extracted color information. Specifically, we specified a square, $L$ pixels in length, centered on $(x_{1}, x_{2})$, that contained a set of pixels surrounding $\bold{x}$ in frame $t$. We chose to capture the hue for each pixel. The set of possible hue values (i.e. the range of hues to which a pixel may be assigned) was partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yielded the $V$ dimensional vector of color counts.

@@

Each of the following synthetic videos consists of a sequence of 200 images (each of size $500$ x $500$ pixels) containing a number of smaller colored squares of different (and potentially time-varying) sizes moving at varied speeds and trajectories over a black background. The synthetic videos contain instances of occlusion (where one or more objects are briefly hidden) and objects with time-varying appearances and behaviors, as these notoriously decrease the accuracy of detection and tracking. After each video was constructed, the extraction procedure described in Section~\ref{sec:dataextraction} (using $L=3$) and inference procedures described in Section~\ref{sec:inference} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices), which are used to determine a sequence of positions and ovals approximating, respectively, the locations and shapes of a tracked object over each frame that it is present in the video (as outlined in Section~\ref{sec:inferencetotrackingresults}).

The first synthetic video experiment aimed to test the ability of the model and inference procedure to maintain the identity of independent objects based on color information alone. Two videos were constructed, both containing a red square (rgb value $[255,0,0]$ and size $20$ x $20$ pixels) and a blue square (rgb value $[0,0,255]$ and size $20$ x $20$ pixels). In both videos, the squares begin at opposite sides of the scene at frame $f=1$ and travel towards each other, arriving at the same location at $f=100$ (where the blue square occludes the red square). The second half of the two videos differ in that both squares in the first video continue in the same direction and end at the other's starting positions at $f=200$, and both squares in the second video reverse directions and end at their initial starting positions at $f=200$. The frame difference extraction yields identical spatial features in both videos; hence, successful tracking depends fully on the incorporation of color information into the model.

Parameters were set to the same values for inference on both videos: $\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right)$, and $q_{0} = (5, \ldots, 5)$. Inference was carried out using the MCMC algorithm (Section~\ref{sec:MCMC}); the MAP sample correctly tracked both colored squares through occlusion in both videos, and is shown in Figure~\ref{fig:synth_one_plot}.

The second synthetic video experiment aimed to test tracking performance under occlusion, object appearance change, and motion change. A video was constructed showing a red square (rgb value $[255,0,0]$), a green square (rgb value $[0,255,0]$), and a blue square (rgb value $[0,0,255]$). The red square was of size $20$ x $20$ pixels, the blue square was of size $15$ x $15$ pixels, and the green square began at size $50$ x $50$ pixels at frame $f=1$, linearly shrinks to $10$ x $10$ pixels at $f=100$, then linearly grows back to $50$ x $50$ pixels by the end of the video, $f=200$. Furthermore, the red and blue squares display the same behavior as in the second video of the first synthetic experiment (they begin at opposite sides of the scene traveling towards each other, cross at the center of the scene at $f=100$, and reverse direction, ending at their initial positions at $f=200$). The green square begins at a point equidistant from the other two squares, intersects with them as they overlap (causing the blue square to occlude the other two), and continues on in a direction at a 20 degree angle from its initial trajectory.

@@

Benchmark video datasets for object tracking and detection have been produced to provide standard scenes on which researchers can compare detection and tracking results. These videos have been primarily produced for surveillance-related workshops---notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)--which provide researchers with video datasets and algorithmic goals on which to focus. Three commonly used benchmark videos from PETS workshops---one used in PETS2000, one in PETS2001, and one used both in PETS2009 and PETS2010---were chosen to demonstrate the method presented in this study. The performance metrics and benchmark datasets allow the methods developed in this paper to be quantitatively compared against other detection and tracking algorithms.

@@

The PETS2000 and PETS2001 video datasets both consist of a small number of humans and vehicles traveling across a parking lot, with video taken from above, emulating what might be recorded by standard outdoor surveillance equipment. The `Test Sequence', a set of images from a monocular, stationary camera, was used from the PETS2000 workshop, and View Two of Dataset 1, also taken via a monocular, stationary camera, was used from the PETS2001 workshop. The MCMC algorithm (described in Section~\ref{sec:MCMC}) was used for inference in these experiments.

>>>>

\begin{abstract}
This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos. It is intended to reduce the need for detection and localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and image qualities. The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be easily and efficiently extracted from the regions in a video that represent objects. This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference in the model to accomplish detection and tracking. This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical characteristics moving over non-uniform backgrounds and through occlusion.
% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
\end{abstract}

@@

We begin by describing characteristics of the extracted data (Section~\ref{sec:dataextraction}), and giving the generalized form of the model (Section~\ref{sec:modeldefinition}). To implement this model, one must specify a data extraction procedure and distributions for representing objects, which may be chosen to allow for arbitrary object tracking or tailored to a specific object type for a given application. In our implementation, we extract data via a basic frame differencing procedure and specify distributions useful for representing arbitrary objects (Section~\ref{sec:modelspecification}). We describe inference algorithms for our model and show how the output of these algorithms can be interpreted as object localization and tracking results (Section~\ref{sec:inference}). Our implementation is demonstrated on multiple synthetic and benchmark datasets. Standard performance metric values are computed to quantify results on the benchmark datasets (Section~\ref{sec:experiments}). We compare the performance metrics of our method with those yielded by specialized detection and tracking algorithms tailored to specific objects in the benchmark datasets. Our results support our hypothesis that approaches combining simple data extraction and a powerful model can perform detection and tracking of arbitrary objects at a level comparable to state-of-the-art, object specific algorithms.

@@

Methods that discern between the foreground and background regions of a video allow for data to be extracted from the areas in each frame where objects reside. Frame differencing and background subtraction are two such methods. Both record locations that exhibit motion relative to the background. Often, background subtraction refers to methods that compare an image containing targets with an image of the background only or with some model of the background that is learned as the video progresses \cite{piccardi2004background}, while frame differencing refers to methods that compare pairs of consecutive images in a video \cite{zhang2001segmentation}. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking \citep{perez_2002}.

@@

Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos. Blob tracking, a basic method for carrying out this goal, has found success in videos where objects are easily isolated from the background and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}. Blob tracking methods, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}. To improve the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}. Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects in videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}. Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions extracted by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and graph based methods that carry out a similar task using spectral clustering \cite{fragkiadaki2011detection}. Another approach uses a Gaussian mixture model to cluster data extracted from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new object tracks.

@@

We desire a data extraction procedure that yields observations of the form
\begin{equation}
\centering
\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
\end{equation}
where each $\bold{x}$ corresponds to a point within an image region where an object (or foreground element)  is believed to reside, $\bold{x}^{s} \in \mathbb{R}^{2}$ denotes the spatial location of this point, $\bold{x}^{c} \in S_{1} \times \ldots \times S_{V}$ denotes some collection of local image features in the vicinity of this point, and $t \in \{1, \ldots, T \}$ denotes the time index.

@@

We also extract features $\bold{x}^{c}$ that capture image information in the vicinity of each extracted pixel. Examples of possible features include color distributions, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations. In principle, we can extract any image features that may be used to characterize the appearance of objects. In our implementation, we choose to extract only color information in the vicinity of each pixel. Incorporating color features allows our method to infer a distribution over color for each detected object; this improves its ability to distinguish between adjacent objects and track objects through occlusion. To add this information, we let $\bold{x}^{c}$ represent a $V$ dimensional discrete distribution over some aspect of color (such as the hue) in the pixel's immediate vicinity. Details on this vector and how it is computed are given in Section~\ref{sec:dataextractioninexperiments}. We refer to the components of this vector as the ``color counts'' of the pixel.

@@

Instead of incorporating the cluster size variable $m_{k,t}$ directly in the GPUDDPM model (as it was in the definition given by \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of new set of variables called deletion variables. We introduce a deletion variable $d_{i,t}$ for each observation $\bold{x}_{i,t}$, which denotes the time at which the observation is removed from its assigned cluster. At each time, cluster sizes $m_{k,t}$ can be reconstructed from all previous assignments and deletion variables by
\begin{equation}
\label{compute_clust_size}
m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
\end{equation}
where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise. Additionally, we can define the deletion variable $d_{i,t}$ to be $d_{i,t} = t + l_{i,t}$, where $l_{i,t}$ can be thought of as the lifetime of an assignment. From the definition of distribution D (given by \eqref{D_distro}), the lifetime can be shown to be distributed geometrically, and can be written as
\begin{equation}
\label{del_rho_form}
l_{i,t} | \rho  \sim  \rho(1 - \rho)^{l_{i,t}}
\end{equation}
where the parameter $\rho$ is the same as that in \eqref{D_distro}.

@@

\subsubsection{Sampling Assignment Variables, $c_{i,t}$}
\label{sec:sample_assignments}

@@

\subsubsection{Sampling Cluster Parameters, $\theta_{k, t}$}

The conjugacy of appearance model and transition kernel distributions allow us to easily sample from the posterior distribution over the cluster parameters, which we can write
\begin{align}
\begin{split}
P(\theta_{k,t} | \ldots ) = & \hspace{1mm} P(\bold{x}_{i,t} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})
\\ & \times P(\theta_{k,t} | z_{k,t,1:M}) \\
 = & \hspace{1mm} \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{k,t}, \Sigma_{k,t} | \boldsymbol{\mu}_{N}, k_{N}, v_{N}, \Lambda_{N}) \\
  & \times \mathcal{D}ir( \bold{p}_{k,t} | \bold{q}_{N})
\end{split}
\end{align}
Where the parameters in the above distribution are given when the observations $\bold{x}_{1:N_{t},t}$, and auxiliary variables $z_{k,t-1:t,1:M}$ for cluster k at time $t-1$ and $t$, are taken to be the ``observations'' for the following Bayesian updates
\begin{eqnarray}
\label{bayesian_update_1}
\kappa_{N} &=& \kappa_{0} + N \\
\label{bayesian_update_2}
\nu_{N} &=& \nu_{0} + N \\
\label{bayesian_update_3}
\boldsymbol{\mu}_{N} &=& \frac{\kappa_{0}}{\kappa_{0}+N} \boldsymbol{\mu}_{0}  +  \frac{N}{\kappa_{0}+N} \overline{\bold{x}}^{s}\\
\Lambda_{N} &=& \Lambda_{0} + S_{\bold{x}^{s}}\\
\label{bayesian_update_4}
\bold{q}_{N} &=& \bold{q}_{0} + \sum_{i=1}^{N} \bold{x}_{i}^{c}
\end{eqnarray}
where $N$ is the number of observations, $\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{x}^{s}$ and $\bold{x}^{c}$ respectively denote the spatial and color features of the observations, and $\overline{\bold{x}}$ and $S_{\bold{x}}$ respectively denote the sample mean and sample covariance for the set of observations $\bold{x}$, defined in \eqref{samplemean} and \eqref{samplecov}.

@@

\subsubsection{Sampling Auxiliary Variables, $z_{k,t,m}$}
\label{sec:sample_aux_vars}

@@

\subsubsection{Sampling Deletion Variables, $d_{i,t}$}
Sampling deletion variables could be performed in a manner similar to how we sample the assignment variables in Section~\ref{sec:sample_assignments}, but this may be computationally expensive due to the large number of possible deletion times. To remedy this, the MH algorithm may again be used to generate samples $d_{i,t}^{*}$ from the posterior distribution over possible deletion times, where we use the proposal distribution
\begin{align}
\begin{split}
l_{i,t}  &\sim  \text{Geo}(\rho)  \\
d_{i,t}^{*}  &= l_{i,t} + t + 1
\end{split}
\end{align}
where $l_{i,t}$ denotes the geometrically distributed ``lifetime'', and we accept or reject this sample using the process described in Section~\ref{sec:sample_aux_vars}.

@@

This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference. SMC inference operates in the orginal GPUDDPM formulation (Section~\ref{sec:gpuddpm}). The algorithm is shown in Algorithm~\ref{alg:SMC}. At each time step $t \in \{ 1, \ldots, T \}$, a number of samples referred to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots,$ $\theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t},$ $\ldots,$ $m_{K_{t},t}$. A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a new set of particles are sampled from the set of weighted particles via a resampling process. Within each time step, Gibbs sampling is used to generate the samples associated with each particle.

@@

\subsubsection{Proposal Distribution for Assignments}
\label{sec:smc_proposal_1}

@@

Performance evaluation metrics, which provide a standardized way of quantifying the success of a detection and tracking procedure on a given video, have started to become consistently used in the past four years. The metrics presented in \cite{kasturi_2008} and used in \cite{ellis_2010, taj_2007, lee_2009} have become well established for evaluating the performance of object detection and tracking in videos and have been adopted by the Video Analysis and Content Extraction (VACE) program and the Classification of Events, Activities, and Relationships (CLEAR) consortium, two large-scale efforts concerned with video tracking and interaction analysis. The two metrics used to quantify the experimental results in this study are known as the Sequence Frame Detection Accuracy (SFDA) and Average Tracking Accuracy (ATA). Details on how these metrics are defined and computed are given in Appendix~\ref{sec:performancemetricdetails}.

@@

The ground-truth authored by the ViPER tool took the form of bounding boxes denoting the spatial position of each object at each time step. Consequentially, to find the spatial overlap between results and ground-truth, which is intrinsic to both metrics, a rectangular bounding box was needed per object per time step from the results of the algorithm. We took the maximal and minimal axially aligned values of the oval inferred by our algorithm (as described in Section~\ref{sec:inferencetotrackingresults}) to be the sides of a representative bounding box for a given object at a given frame.

@@

Frame differencing was used in all experiments to identify pixels exhibiting motion. 
For each pixel $\bold{x} = ( x_{1},$ $x_{2},$ $t )$ recorded during frame differencing, we also extracted color information. Specifically, we specified a square, $L$ pixels in length, centered on $(x_{1}, x_{2})$, that contained a set of pixels surrounding $\bold{x}$ in frame $t$. We chose to capture the hue for each pixel. The set of possible hue values (i.e. the range of hues to which a pixel may be assigned) was partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yielded the $V$ dimensional vector of color counts. For all experiments, we chose $V=10$.

@@

Each of the following synthetic videos consists of a sequence of 200 images (each of size $500$ $\times$ $500$ pixels) containing a number of smaller colored squares of different (and potentially time-varying) sizes moving at varied speeds and trajectories over a black background. The synthetic videos contain instances of occlusion (where one or more objects are briefly hidden) and objects with time-varying appearances and behaviors, as these notoriously decrease the accuracy of detection and tracking. After each video was constructed, the extraction procedure described in Section~\ref{sec:dataextractioninexperiments} (using $L=3$) and inference procedures described in Section~\ref{sec:inference} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices), which are used to determine a sequence of positions and ovals approximating, respectively, the locations and shapes of a tracked object over each frame that it is present in the video (as outlined in Section~\ref{sec:inferencetotrackingresults}).

The first synthetic video experiment aimed to test the ability of the model and inference procedure to maintain the identity of independent objects based on color information alone. Two videos were constructed, both containing a red square (rgb value $[255,0,0]$ and size $20$ $\times$ $20$ pixels) and a blue square (rgb value $[0,0,255]$ and size $20$ $\times$ $20$ pixels). In both videos, the squares begin at opposite sides of the scene at frame $f=1$ and travel towards each other, arriving at the same location at $f=100$ (where the blue square occludes the red square). The second half of the two videos differ in that both squares in the first video continue in the same direction and end at the other's starting position at $f=200$, and both squares in the second video reverse directions and end at their initial starting positions at $f=200$. The frame difference extraction yields identical spatial features in both videos; hence, successful tracking depends fully on the incorporation of color information into the model.

Parameters were set to the same values for inference on both videos: $\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right)$, and $\bold{q}_{0} = (5, \ldots, 5)$. Inference was carried out using the MCMC algorithm (Section~\ref{sec:MCMC}); the MAP sample correctly tracked both colored squares through occlusion in both videos, and is shown in Figure~\ref{fig:synth_one_plot}.

The second synthetic video experiment aimed to test tracking performance under occlusion, object appearance change, and motion change. A video was constructed showing a red square (rgb value $[255,0,0]$), a green square (rgb value $[0,255,0]$), and a blue square (rgb value $[0,0,255]$). The red square was of size $20$ $\times$ $20$ pixels, the blue square was of size $15$ $\times$ $15$ pixels, and the green square began at size $50$ $\times$ $50$ pixels at frame $f=1$, linearly shrinks to $10$ $\times$ $10$ pixels at $f=100$, then linearly grows back to $50$ $\times$ $50$ pixels by the end of the video, $f=200$. Furthermore, the red and blue squares display the same behavior as in the second video of the first synthetic experiment (they begin at opposite sides of the scene traveling towards each other, cross at the center of the scene at $f=100$, and reverse direction, ending at their initial positions at $f=200$). The green square begins at a point equidistant from the other two squares, intersects with them as they overlap (causing the blue square to occlude the other two), and continues on in a direction at a 20 degree angle from its initial trajectory.

@@

Benchmark video datasets for object tracking and detection have been produced to provide standard scenes on which researchers can compare detection and tracking results. These videos have been primarily produced for surveillance-related workshops---notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)---which provides researchers with video datasets and algorithmic goals on which to focus. Three commonly used benchmark videos from PETS workshops---one used in PETS2000, one in PETS2001, and one used both in PETS2009 and PETS2010---were chosen to demonstrate the method presented in this study. The performance metrics and benchmark datasets allow the methods developed in this paper to be quantitatively compared against other detection and tracking algorithms.

@@

The PETS2000 and PETS2001 video datasets both consist of a small number of humans and vehicles traveling across a parking lot, with video taken from above, emulating what might be recorded by standard outdoor surveillance equipment. The ``Test Sequence'', a set of images from a monocular, stationary camera, was used from the PETS2000 workshop, and ``View Two of Dataset 1", also taken via a monocular, stationary camera, was used from the PETS2001 workshop. The MCMC algorithm (described in Section~\ref{sec:MCMC}) was used for inference in these experiments.
